---
title: "Regression under differential privacy"
author: "LJ"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: yeti
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,  message=FALSE, warning=FALSE}
library(readxl)
library(data.table)
library(xlsx)
library(pROC)
library(ROCR)
library(tidyverse)
library(caret)
library(ggpubr)
library(rms)
library(dplyr)
library(ggplot2)
library(DT)
library(tableone)
library(glmnet)
library(ncvreg)
library(MASS)
library(reshape2)
library(Hmisc)
library(sjPlot) # beautiful summaries
library(sjmisc)
library(sjlabelled)


ljtheme <- theme_bw() + theme(panel.border = element_rect(color = "black", size = 2),
                                 # Remove panel grid lines
                                panel.grid.major = element_blank(),
                                panel.grid.minor = element_blank())
```

#### Regression Analysis on differentially private synthetic datasets

Data release without disclosure of personal information is increasingly challenging with recent
technological advancements. Our presence in multiple databases (e.g. social media sites) increases
the prospect of participant re-identification in de-identified dataset through linkage attacks. Differ-
ential privacy (DP) provides a formal guarantee that an individual participant’s data would remain
private. However, the utility of synthetic datasets generated through DP mechanisms are sometimes
undesirable. Pooled (or aggregate) synthetic data has proven a reliable substitute for individual
data without the formal privacy guarantees provided by DP mechanisms. To bridge this gap, we
propose a hybrid mechanism, a pooled differentially private algorithm (PoolDiv), that utilizes a relax
(, δ) privacy budget and the data aggregation techniques of pooling to generate synthetic data. We
conduct extensive simulations to compare the performance of least square estimation on synthetic data
generated from (1) traditional DP mechanisms (i.e. randomized response mechanism and histogram
perturbation algorithm), (2) pooling data synthesis mechanism, and (3) PoolDiv. We assess the
performance of the mechanism on simulated datasets and on real data.

### simulate data here

```{r,eval=T, echo=T, fig.align="center", fig.height=5, fig.width=5, warning=FALSE, message=F}

# simulate the data
sim_lsdata<-function(n=1000, p=4, sigma.star=1/2){
  beta.star=rnorm(p)
  X=cbind(1, matrix(rnorm(n*(p-1)), nrow=n, ncol=(p-1))) 
  y=X%*%beta.star + sigma.star*rnorm(n)
  data=data.frame(cbind(y,X)); 
  colnames(data)<-c("y",paste0("X",0:(p-1),sep = ""))
  dat<-data[,-2]
  return(data.frame(dat) )
}

simdat = sim_lsdata(n=1000, p=4, sigma.star=1/2)


xbar=apply(simdat, 2, mean); 
simdat=scale(simdat, center=xbar, scale=FALSE)
```


#### Laplace mechanism

```{r,eval=T, echo=T, fig.align="center", fig.height=3.4, fig.width=6, warning=FALSE, message=F}
library(rmutil)
lap_values = rlaplace(n = 1000,m = 0,s = 0.1)

N = 1000
r5 <- rlaplace(n = N,m = 0,s = 5)
r4 <- rlaplace(n = N,m = 0,s = 4)
r3 =  rlaplace(n = N,m = 0,s = 3)
r2 <- rlaplace(n = N,m = 0,s = 2)
r1 <- rlaplace(n = N,m = 0,s = 1)

data_list<-list(eps5 = r5,eps4 = r4,eps3 = r3,eps2 = r2, eps1 = r1)

data<- melt(data_list)

ggplot(data,aes(x=value, col=L1)) + geom_density(alpha=0.25,size=0.8) + ljtheme  + 
  scale_color_discrete(name = "Privacy budget",
                 limits = c("eps1","eps2","eps3","eps4","eps5"),
                 labels = c(paste(expression(epsilon),"=",1),
                            paste(expression(epsilon),"=",2),
                            paste(expression(epsilon),"=",3),
                            paste(expression(epsilon),"=",4),
                            paste(expression(epsilon),"=",5))) + xlim(c(-30,30)) +
  theme(legend.position = c(0.8, 0.6),
        legend.background = element_rect(fill="lightgray", 
                                  size=0.5, linetype="solid"))

```
remark: laplace noise addition for continuous variable

#### random mechanisns

randomized response mechanism

```{r,eval=T, echo=T, fig.align="center", fig.height=3.4, fig.width=6, warning=FALSE, message=F}

alpha = beta = 0.5
# the randomized mechanism
random_mechanism = function(d, alpha=0.5, beta=0.5){
  foo = function(d){
  if(runif(1,0,1)<alpha){
    return(d)
    } 
  else if(runif(1,0,1)<beta){
      return(runif(n = 1,min = -5,max = 5))
      }
  else{ return(d)}
  }
  return(apply(as.matrix(d), 1, foo))
}

N = 10^4
#data = rnorm(N,mean = 0,n = 25)
data = simdat[,1]

r2 <- random_mechanism(data,alpha = 0.9,beta = 0.75)
r3 <- random_mechanism(data,alpha = 0.8,beta = 0.75)
r4 =  random_mechanism(data,alpha = 0.5,beta = 0.75)
r5 <- random_mechanism(data,alpha = 0.3,beta = 0.75)
r1 <- data

data_list<-list(r5 = r5,r4 = r4,r3 = r3,r2 = r2, r1 = r1)

data<- melt(data_list)

ggplot(data,aes(x=value, col=L1)) + geom_density(alpha=0.25,size=0.9) + ljtheme  + 
  scale_color_discrete(name = "Randomization",
                 limits = c("r1","r2","r3","r4","r5"),
                 labels = c("non-random",
                            paste(expression(a),"=",0.9),
                            paste(expression(a),"=",0.8),
                            paste(expression(a),"=",0.5),
                            paste(expression(a),"=",0.3)))+
  theme(legend.position = c(0.87, 0.67),
        legend.background = element_rect(fill="lightgray", linetype="solid"))
```
remark: random mechanism sampling

#### exponential mechanism

```{r,eval=T, echo=T, fig.align="center", fig.height=3.4, fig.width=6, warning=FALSE, message=F}
library(rmutil)
gaus_values = rnorm(n = 1000,mean = 0,sd = 0.5)

N = 1000
r1 <- rnorm(n = 1000,mean = 0,sd = 1)
r2 <- rnorm(n = 1000,mean = 0,sd = 2)
r3 =  rnorm(n = 1000,mean = 0,sd = 3)
r4 <- rnorm(n = 1000,mean = 0,sd = 4)
r5 <- rnorm(n = 1000,mean = 0,sd = 5)

data_list<-list(eps5 = r5,eps4 = r4,eps3 = r3,eps2 = r2, eps1 = r1)

data<- melt(data_list)
ggplot(data,aes(x=value, col=L1)) + geom_density(alpha=0.25,size=0.9) + ljtheme  +
    scale_color_discrete(name = "noice level",
                 limits = c("eps1","eps2","eps3","eps4","eps5"),
                 labels = c(paste(expression(dispersion),"=",1),
                            paste(expression(dispersion),"=",2),
                            paste(expression(dispersion),"=",3),
                            paste(expression(dispersion),"=",4),
                            paste(expression(dispersion),"=",5))) +
  theme(legend.position = c(0.8, 0.6),
        legend.background = element_rect(fill="lightgray", 
                                  size=0.5, linetype="solid"))

```
remark: see 064XXX.pdf


### Generating synthetic data


#### Algorithm 1: 
- randomized function k for a binary dataset

We use here a randomized algorithm to generate M multiple datasets each of budget $\epsilon /M$.


```{r,eval=T, echo=T, fig.align="center", fig.height=3.5, fig.width=4, warning=FALSE, message=F}

# Toy binary data
ybinary<-ifelse(simdat[,1]>0.5,0,1)
# assume a binomial likelihood for the data
# sufficient stats
 ybar = mean(ybinary)
 
 # sample from binomial
 data = ybinary
 syn_private_binom<-function(data,alp1=0.9,alp2=1.75,n_out=500, epsil = 8){
   if (min(alp1,alp2)<n_out/(exp(epsil)-1)) {
     print("differential privacy condition not satisfied")
   }
   n = length(data)
   x = sum(data)
   Q =  seq(0, 1, length = n_out)
   pbar = pbeta(q = Q,shape1 = alp1+x, shape2 = alp2+n-x)
   result = rbinom(n = length(pbar),size = 1,prob = pbar)
   return(ifelse(result>0,0,1))
 }
 
synval = syn_private_binom(data)
#chisq.test(data,syntheticval)
dat = data.frame(val=as.factor(c(synval,data)),
                 category = c(rep("New",length(synval)),
                              rep("Private",length(data))))

#str(dat)
vars = c("val")
Table1<-CreateTableOne(vars = vars, strata = "category",
                       data = dat, test = TRUE)
print(Table1, smd=F,showAllLevels = TRUE,noSpaces = TRUE)
levels(dat$category)<-c("Synthetic data","Private data")

cat("Toy example for beta-binomial mechanism")
ggplot(dat, aes(category, ..count..)) + geom_bar(aes(fill = val), position = "dodge") + ljtheme + 
    scale_fill_discrete(name = "category") +
  theme(legend.position = c(0.2, 0.8),
        legend.background = element_rect(fill="lightgray", 
                                  size=0.5, linetype="solid"))

```

Remark: synthetic binary distribution generation


#### Algorithm II
Histogram Perturbation for contingency tables. The algorithm is based on the contingency table representation of data

```{r,eval=T, echo=T, fig.align="center", fig.height=5, fig.width=6, warning=FALSE, message=F}

# Toy contingency table data
JobSat <- matrix(c(1,2,1,0, 3,3,6,1, 10,10,14,9, 6,7,12,11), 4, 4)
dimnames(JobSat) = list(income=c("< 15k", "15-25k", "25-40k", "> 40k"),
                        satisfaction=c("VeryD", "LittleD", "ModerateS", "VeryS"))
#as.data.frame(JobSat)
JobSat <- as.table(JobSat)
#str(JobSat)
cat("private contingency table dataset")
JobSat

# For spliting continuous variables into quantiles
qvals = function(X,P=seq(0,1,length.out = 20)){
  val = table(cut(X, breaks = quantile(X,probs = P)))
  return( as.matrix(val)[,1] )
}
#apply(simdat, 2, qvals)
#dat = as.data.frame(apply(simdat, 2, qvals))

# To measure the sensitivity of database
Senseval = function(dat){ # privacy sensitivity
  return(abs(max(dat)-min(dat)))
}



dat<-as.matrix(as.vector(JobSat))
perturb_hist<-function(dat,epsil = 0.005){ # def histogram perturbation function
  sense = Senseval(dat) 
  noise = replicate(n = length(dat),
                  expr = rlaplace(m = 0,s = epsil/sense))
  di = dat/sum(dat) + noise
  di_bar = ifelse(di<0,0,di)
  pi_bar = di_bar/sum(di_bar)
  return(pi_bar )
}

temp = matrix(as.vector(apply(dat, 2, perturb_hist)),nrow = 1)

#sample from multinomial
Xreturn = rMultinom(probs = temp,m = 16)

syntheticdat = as.table(matrix(Xreturn,nrow = 4))
rownames(syntheticdat)<-rownames(JobSat)
colnames(syntheticdat)<-colnames(JobSat)

syntheticdat = as.data.frame(syntheticdat)
returnprivate<-as.data.frame(JobSat)

colnames(syntheticdat)<-colnames(returnprivate)
all_dat<-rbind(returnprivate,syntheticdat)
all_dat$datatype<-c(rep("private",length(returnprivate$Freq)),
                    rep("synthetic",length(syntheticdat$Freq)))

all_data_flat = all_dat %>% 
  slice(rep(1:n(), Freq)) 

df = all_dat
cat("Toy example with count data")
ggplot(df, aes(income, Freq)) +
  geom_bar(stat = "identity", position = position_dodge(0.8),width = 0.7,aes(fill = satisfaction))+
  ylab("Count (N)") + xlab("Income") + #coord_flip() +
  facet_grid(datatype ~., scales = "free")  + theme(axis.text.x = element_text(vjust = +.65,angle = 45, size = 0.15)) +
  labs(subtitle="Private data vs synthetic data")+ 
  scale_fill_hue(name = "Income satisfaction") + ljtheme +
    theme(legend.position = "bottom",
        legend.background = element_rect(fill="lightgray", 
                                  size=0.5, linetype="solid"))

```



#### Algorithm III

We use here the Multinomial-Dirichlet Synthesizer. The algorithm is specified in the manuscript.

Remark: Not included in this assessment



### Pooled synthetic data
```{r,eval=T, echo=T, fig.align="center", fig.height=4, fig.width=4, warning=FALSE, message=F}

# Function to create synthetic pool data
pool<-function(poolsize=2,data){
  G = poolsize;
  n = dim(data)[1]
  pools<-n/G
  pooldata<-data.frame(matrix(0,nrow = pools,ncol = NCOL(data)))
  colnames(pooldata)<-colnames(data)
  if (poolsize==2) {
    pooldata<-data[seq(1,n-1,G),]+data[seq(2,n,G),]
  }
  if (poolsize==4) {
    
    pooldata<-parse(text = paste(paste(paste("data[seq(",1:G,",n -G +",1:G,",G),] + "),
                               collapse = ""),"0") )
    pooldata = eval(pooldata)
    }
  
  if (poolsize==6) {
    
    pooldata<-parse(text = paste(paste(paste("data[seq(",1:G,",n -G +",1:G,",G),] + "),
                               collapse = ""),"0") )
    pooldata = eval(pooldata)
  }
  
  if (poolsize==10) {
    pooldata<-parse(text = paste(paste(paste("data[seq(",1:G,",n -G +",1:G,",G),] + "),
                               collapse = ""),"0") )
    pooldata = eval(pooldata)
  }
  
  return(pooldata)
}


poolsize=2
pool2  = pool(poolsize=2,simdat)

```


### All the datasets for comparison

- simulated data
- differential private synthetic data using randomized algorithm
- differential private synthetic data using histogram perturbation algorithm
- pooled synthetic data
- pooled differentially private synthetic data

```{r,eval=T, echo=T, fig.align="center", fig.height=4, fig.width=4, warning=FALSE, message=F}
simdat<-as.data.frame(simdat)
train.control <- trainControl(method = "cv", number = 10, repeats = 5)

simulated<-simdat
diff_index<-random_mechanism(simdat[,1],alpha = 0.75,beta = 0.6)
diff_randomized<-simdat
diff_randomized$y<-diff_index; randomized_keep<-diff_randomized
## histogram perturbation, used all_dat
# laplace noise
diff_laplace <-simdat
laplace_mech<-function(data,epsil=3.5){
  return(data + rlaplace(n = length(data),m = 0,s = epsil/(Senseval(data)))
         )
}
diff_laplace<-apply(diff_laplace,2,laplace_mech)

diff_hist<-all_data_flat
pool2  = pool(poolsize=2,simdat)
pool4  = pool(poolsize=4,simdat)
pool6  = pool(poolsize=6,simdat)


lm_private <- train(y ~., data = simulated, method = "lm",
               trControl = train.control)
cat("Resampled RMSE: Private data")
lm_private


lm_random <- train(y ~., data = diff_randomized, method = "lm",
               trControl = train.control)
cat("Resampled RMSE: Randomized response")
lm_random


lm_laplace <- train(y ~., data = diff_laplace, method = "lm",
               trControl = train.control)
cat("Resampled RMSE: Laplace noise perturbation")
lm_laplace


lm_pool2 <- train(y ~., data = pool2, method = "lm",
               trControl = train.control)
cat("Resampled RMSE: Pooled synthetic (g=2)")
lm_pool2


lm_pool4 <- train(y ~., data = pool4, method = "lm",
               trControl = train.control)
cat("Resampled RMSE: Pooled synthetic (g=4)")
lm_pool4

lm_pool6 <- train(y ~., data = pool6, method = "lm",
               trControl = train.control)
cat("Resampled RMSE: Pooled synthetic (g=6)")
lm_pool6



```


### Compile the result

```{r,eval=T, echo=T, fig.align="center", fig.height=5, fig.width=7, warning=FALSE, message=F}
all_rslt<-rbind(
lm_private$results[2:4],
lm_random$results[2:4],
lm_laplace$results[2:4],
lm_pool2$results[2:4],
lm_pool4$results[2:4],
lm_pool6$results[2:4]
)
all_rslt$Data<-c("Private","Randomized","Laplace","Pool2","Pool4","Pool6")

all_rslt_long<-gather(all_rslt,"Metric","Value",1:3)

# plot all the results separated by the metric used fo rthe analysis 
df = all_rslt_long
ggplot(df, aes(Data, Value)) +
  #geom_bar(stat = "identity",size=0.5, position = position_dodge(0.8),width = 0.7,aes(fill=Data)) +
  geom_point(size=3) + 
  geom_segment(aes(x=Data, 
                   xend=Data, 
                   y=0, 
                   yend=Value)) + 
  ylab("Measure") + xlab("Dataset") + coord_flip() +
  facet_grid(Metric ~., scales = "free")  + 
  theme(axis.text.x = element_text(vjust = +.65,angle = 45, size = 0.15)) +
  labs(subtitle="Regression Analysis on datasets")+ 
  scale_fill_hue(name = "") + ljtheme +
    theme(legend.position = "none",
        legend.background = element_rect(fill="lightgray", 
                                  size=0.5, linetype="solid"))


```



```{r,eval=T, echo=T, fig.align="center", fig.height=5, fig.width=7, warning=FALSE, message=F}
lm_private <- lm(y ~., data = simulated)
lm_random <- lm(y ~., data = diff_randomized)
lm_laplace <- lm(y ~., data = as.data.frame(diff_laplace) )
lm_pool2 <- lm(y ~., data = pool2)
lm_pool4 <- lm(y ~., data = pool4)
lm_pool6 <- lm(y ~., data = pool6)

vals=tab_model(lm_private,lm_random,lm_laplace,lm_pool2,lm_pool4,lm_pool6,show.p = F,auto.label = FALSE)
vals
```


### Pooled epsilon differential data
```{r,eval=T, echo=T, fig.align="center", fig.height=5, fig.width=7, warning=FALSE, message=F}

divpool2  = pool(poolsize=2,randomized_keep)
divpool4  = pool(poolsize=4,randomized_keep)
divpool6  = pool(poolsize=6,randomized_keep)
divpool10  = pool(poolsize=10,randomized_keep)

lm_pool2 <- lm(y ~., data = divpool2)
lm_pool4 <- lm(y ~., data = divpool4)
lm_pool6 <- lm(y ~., data = divpool6)
lm_pool10 <- lm(y ~., data = divpool10)

tab_model(lm_pool2,lm_pool4,lm_pool6,lm_pool10)
```


### Real data example
- Here we will look at a high dimensional data set with $p>>n$. We will use the lymphoma dataset used by Shipp et al. (2002). It consist of 58 individuals and 6817 gene expression profiles. The outcome is disease status (yes/no).

```{r,eval=T, echo=T, fig.align="center", fig.height=5, fig.width=7, warning=FALSE, message=F}
library(datamicroarray)

# describe_data()
data('shipp', package = 'datamicroarray')
dim(shipp$x);table(shipp$y) # Always run directly in the console

df<-scale(shipp$x[,1:100],center = T)
y<-shipp$y

library("d3heatmap")
#d3heatmap(scale(df[,sample(1:100,25)]), colors = "RdYlBu")
#heatmap(df[,sample(1:100,30)])

library(gplots)
cat("Private data")
heatmap.2(df, scale = "none", col = bluered(100), 
          trace = "none", density.info = "none",main = "[Private]")

cat("randomized")
# randomized private data
diff_index<-random_mechanism(y,alpha = 0.75,beta = 0.6)
diff_randomized<-as.matrix(df[which(y%in%c("DLBCL","FL")),])
heatmap.2(diff_randomized, scale = "none", col = bluered(100), 
          trace = "none", density.info = "none",main = "[Randomized]")

##3 laplace noise addition
cat("laplace data")
diff_laplace<-apply(df,2,laplace_mech)
heatmap.2(diff_laplace, scale = "none", col = bluered(100), 
          trace = "none", density.info = "none",main = "[Laplace]")

# pool2

pool2  = pool(poolsize=2,df)
pool4  = pool(poolsize=4,df)
pool6  = pool(poolsize=6,df)
cat("Pool 2")
heatmap.2(pool2, scale = "none", col = bluered(100), 
          trace = "none", density.info = "none",main = "[Pool-2]")
cat("Pool 4")
heatmap.2(pool4, scale = "none", col = bluered(100), 
          trace = "none", density.info = "none",main = "[Pool-4]")

cat("Pool 6")
heatmap.2(pool6, scale = "none", col = bluered(100), 
          trace = "none", density.info = "none",main = "[Pool-6]")

```



